Solar wind predictions for the Parker Solar Probe orbit

### 2017-10-12 paperMVVB_referee/referee_report_responses.txt

- adopted
+ highlighted
+ can be adopted

Next paragraph, last sentence, suggest: the perihelion distance to be reduced to a minimum of 9.86 Rs. Reverse figures 10 and 11, since figure 11 is referenced first, or just remove the mention of Figure 11 here, which isn't necessary; the statement can be understood without looking at the figure.

- adopted
+ highlighted
+ can be adopted

- switched order of figures 10 and 11
- added top and bottom pointers to both references of now figure 10
+ highlighted
+ manage fig references

@@

Last line of page: How are the ranges of median values for different SSN shown in the figure? I'm guessing that it's the thickness of the shaded regions, but this could be stated specifically to guide the reader in interpreting the figure.

- added remark about that
+ highlighted
+ explain thickness of shaded region

- added further explanation in figure caption

@@

- adopted
+ highlighted
+ can be adopted

- ordered the references Sheeley 1997 and Wang 2000 by year

Line 3: But do the extreme values really occur for min and max SSN? E.g. the highest speeds could be at intermediate SSN? How can a time lag be “noted” in this figure which there is no obvious time dependence?

+ yes, in average, see figure 5

- correct, removed remark about time lag from figure caption
+ remove remark about time lag from caption

- added column to table 2: SSN time lag

@@

Line 5, suggest: mentioned changing shape of the frequency distribution with heliocentric distance

- adopted
+ highlighted
+ can be adopted

Line 6: But the largest difference appears to be at ~1 AU, so isn't there also a problem with the dipole/quadrapole model matching the observed fields at 1 AU?

+ highlighted
+ DQCS model assumes 3.1 nT at 1 au coming from CHs at +-60° latitudes (excluding field from equatorial slow wind belt)

@@

Line 7: “Whereas”: Combine the sentences; whereas isn’t used at the beginning of a sentence. Since the peak speeds are on the declining phase of the cycle, wouldn't the peak predicted speeds occur for some intermediate sunspot number, not at SSN=200?

- combined
+ highlighted
+ combine

- correct, it is a shifted SSN. added value of time lag (at each solar wind parameter)
+ explain

@@

- adopted
+ highlighted
+ can be adopted

- replaced solar radii with Rs in discussion
+ search text for solar radii and replace with Rs

Paragraph 2, line 3, suggest: Differences from the results of these other studies suggest that below about 20 Rs, PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur.

- adopted with addition
+ highlighted
+ can be adopted, but
+ pay attention to wording

@@

Last line on page, suggest: because the solar wind is believed to be accelerated up to the height of the Alfvenic critical surface, which is predicted to lie on average around 17 Rs, with a range of between 15 Rs at solar minimum and 30 Rs at solar maximum.

- adopted with addition
+ highlighted
+ can be adopted, but
+ check wording

### 2017-10-12 paperMVVB/sections/content.tex
#3 worked on referee comments; made changes bold face

%\subsection{SSN fitting}
\textbf{Next we create solar activity dependent analytical representations of the solar wind frequency distributions. This is achieved by shifting the median positions of the lognormal distributions as a linear function of SSN. To enable these shifts }, we add a linear SSN dependency to the median
\begin{align}
	x_\text{med}(ssn) &= a_\text{med} \cdot ssn + b_\text{med}\,,	\label{eq:median_with_ssn}
\end{align}
using a factor to the SSN $a_\text{med}$ with a baseline $b_\text{med}$. We relate the mean with a scaling factor to the median to transfer its SSN dependency:
\begin{align}
	x_\text{avg}(ssn) &= (1 + a_\text{avg}) \cdot x_\text{med}(ssn)\,.	\label{eq:mean_with_ssn}
\end{align}
\textbf{These relations, substituted into the lognormal function (\ref{eq:single_lognormal_fit_function}), lead to a new SSN-dependent function $W'(x,ssn)$. This function} is then fitted to the yearly data\textbf{, using the yearly SSN as input parameter.} \textbf{The SSN is offset with the individual time lags for each parameter determined before, to profit from the higher correlation.} The three resulting fit coefficients ($a_\text{med}, b_\text{med}$ and $a_\text{avg}$) are presented in Table~\ref{tab:ssn_fit_parameters}.
\begin{table*}
	\caption{Resulting fit coefficients from the OMNI data fitting with \textbf{the linear SSN dependencies (\ref{eq:median_with_ssn}) and (\ref{eq:mean_with_ssn})}. For the velocity the fit parameters from the double lognormal fit \textbf{(\ref{eq:double_lognormal_fit_function})} and their balancing function \textbf{(\ref{eq:balance_with_ssn})} are given. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value.}}
	\label{tab:ssn_fit_parameters}
	\centering
	\begin{tabular}{l@{} c@{}
		S[table-format = 1.3(2)e+1]
		S[table-format = 1.4(2)]
		S[table-format = 1.3(2)e+1]
		S[table-format = +1.3(2)e+1]
		c
		}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{2}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean}	&\multicolumn{2}{c}{Balance}\\
		\cline{3-4}\cline{6-7}
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{SSN factor $a_\text{med}$}	&\multicolumn{1}{c}{Baseline $b_\text{med}$}	&\multicolumn{1}{c}{Scaling factor $a_\text{avg}$}	&\multicolumn{1}{c}{SSN factor $c_a$}	&\multicolumn{1}{c}{Baseline $c_b$}\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&1.309(19)e-2	&4.285(17)	&8.786(78)e-2	&\multicolumn{1}{c}{--}	&--\\
%		\multicolumn{2}{l}{Velocity}	&2.514(93)e-3	&3.8521(90)	&2.392(26)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Density}	&3.81(25)e-3	&4.495(26)	&3.050(27)e-1	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Temperature}	&1.974(26)e-2	&5.729(19)	&6.541(28)e-1	&\multicolumn{1}{c}{--}	&--\\
		\hline
		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W'_1$}	&\multicolumn{1}{c}{--}	&3.633(12)	&1.008(37)e-2	&\multirow{2}{*}{\tablenum{-1.799(95)e-3}}	&\multirow{2}{*}{0.638(32)}\\
			&\multicolumn{1}{c}{$W'_2$}	&\multicolumn{1}{c}{--}	&4.831(81)	&2.31(20)e-2	&	&\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{\hl{In units of } \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

@@

Next we retrieve the frequency distributions of the four parameters in distance bins of \SI{0.01}{\au},  choosing the same resolution as for the OMNI data analyzed in Sect.~\ref{sec:frequency_distribution}---the distributions are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}.

@@

For simplification, as mentioned before, we treat the exponents of the median and mean fit functions as being identical\textbf{, using one fit parameter for both.} Implementing the power law distance dependency~(\ref{eq:power_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get the fit parameters ($d'_\text{med}$, $d'_\text{avg}$ and their common exponent $e'$). Again, we use the double lognormal function~(\ref{eq:double_lognormal_fit_function}) for the velocity distribution fit, resulting in $W''_\text{II}(x,r)$. The additional fit parameters are the balancing parameter $c'$ and for the second lognormal part $d'_\text{med,2}$ and $d'_\text{avg,2}$. The resulting fit coefficients for the four solar wind parameters are presented in Table~\ref{tab:extrapolation_model_fit_parameters}.

@@

The frequency distribution data for the four solar wind parameters with respect to the radial distance from the Sun \hl{are plotted in } Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}, \hl{together with their power law lognormal fits and the double lognormal fit for the velocity with their median values. } 
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Frequency distributions of the four solar wind parameters with respect to the solar distance. Plotted are the binned Helios data and the power law lognormal fit models with their median values (white lines). The double lognormal model is used for the velocity, its slow and fast parts are indicated by dotted lines.}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}
The model’s magnetic field strength is broader around values of \SI{40}{nT} at the lower distance boundary than the data's frequency distribution implies. This behavior is expected because of the distance independent shape approximation applied. The velocity and temperature models’ upper values generally show a higher abundance than the actual data, see also zoom boxes in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}. \textbf{The increasing high velocity tail with distance} comes from using the same exponent for both slow and fast components. This effect is not seen in the data, more specifically, not only the slowest wind but also the fastest wind is expected to converge to more average speeds \citep{Sanchez-Diaz2016}.
%\citep{Sanchez-Diaz2016} p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction)

@@

%\subsection{Near-Sun extrapolation for PSP orbit}
%PSP orbit
Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching already 93~days after launch in November 2018 a first perihelion with a distance of \SI{0.16}{\au}. Seven additional Venus flybys allow \textbf{the perihelion distance to be reduced} to a minimum of \SI{9.86}{\Rs} \citep{Fox2015} as plotted in \hl{Fig. } ~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.

We extrapolate the derived empirical solar wind model (\ref{eq:general_sw_model}) to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}.

@@

	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. \textbf{The models obtained from Helios and OMNI measurements} are extrapolated to the PSP region---for the range between the extreme cases of solar minimum and maximum, i.e., $0 \le ssn \le 200$ (shaded areas). Note that there is a time lag to the SSN depending on the solar wind parameter. Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \citet{Wang2000} and \citet{Sheeley1997}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}

@@

@@

The magnetic field strength is found to increase from median values of about \SI{43}{\nT} at \SI{0.25}{\au} to \SI{715}{\nT} at \SI{0.046}{\au} \hl{for a SSN of 0. Taking a SSN of 200 } increases the values to \SI{69}{\nT} and \SI{1152}{\nT}. Our extrapolation results are slightly flatter than those derived from the analytical magnetic field model by \citet{Banaszkiewicz1998}, who constructed an analytic dipole plus quadrupole plus current sheet (DQCS) model for solar minimum.

@@

The average velocity is found to decrease from \SI{340}{\km\per\s} at \SI{0.25}{\au} to about \SI{290}{\km\per\s} at \SI{0.046}{\au} for a SSN of 0\textbf{, whereas} using a SSN of 200 it decreases from \SI{390}{\km\per\s} to \SI{330}{\km\per\s}. Comparing the results with the measurements by \citet{Sheeley1997} and \citet{Wang2000} shows an overestimation in our extrapolated slow solar wind velocity values for distances below about \SI{20}{\Rs}. They used LASCO coronagraph observations to track moving coronal features (blobs) in the distance range \SIrange{2}{30}{\Rs} to determine speed profiles and sources of the slow solar wind and they derived temperature and sonic point values for slow solar wind with the isothermal expansion model from \citet{Parker1958}. Therefore, it generally can be expected that PSP will encounter a slower solar wind environment close to the Sun than our model estimates\textbf{. Thus} PSP will measure solar wind acceleration processes \citep{McComas2008}, maybe even still at \SI{30}{\Rs} as the study by \citet{Sheeley1997} suggests.

@@

The proton density increases from about \SI{84}{\per\cm\cubed} at  \SI{0.25}{\au} to about \SI{3018}{\per\cm\cubed} at \SI{0.046}{\au} for a SSN of 0. Being almost independent of the SSN the values for a SSN of 200 are only \SI{17}{\%} larger. The results are in good agreement with those of \citet{Leblanc1998}, who derived an electron density model from type~III radio burst observations. Their model shows that the density distance dependency scales with $r^{-2}$ and steepens just below \SI{10}{\Rs} with $r^{-6}$. \hl{For the comparison we assumed a solar wind helium abundance of } \SI{5}{\%}.

@@

The extrapolated proton temperature increases from about \SI{260000}{\K} at \SI{0.25}{\au} to about \SI{1690000}{\K} at \SI{0.046}{\au} for a SSN of 0 and from \SI{440000}{\K} to \SI{2860000}{\K} for a SSN of 200. Knowing that near-Sun coronal temperatures are in the range of \SIrange{2}{3}{\mega\K} \citep{Billings1959,Liebenberg1975}, the model \textbf{overestimates} the extrapolated temperatures at the PSP perihelion distance.

%\subsection{SSN prediction for PSP mission time}
%SSN prediction
\hl{For SSN short-term predictions } several sources are available. The SIDC provides 12-month SSN forecasts\footnote{\url{http://sidc.be/silso/forecasts}} obtained from different methods (e.g., Kalman filter combined method). The SSN prediction of NOAA's Space Weather Prediction Center (SWPC) follows for the time period until end of 2019 a consensus of the Solar~Cycle~24 Prediction~Panel\footnote{\url{http://www.swpc.noaa.gov/products/solar-cycle-progression}}.
The SSN for PSP's first perihelion will be small---certainly below 20, whereas \textbf{the SSN during the closest perihelia, which will commence at the end of 2024 at the likely maximum phase of cycle~25, cannot be predicted at this time.}
\hl{For the prediction of the next solar cycle we simply assume a pattern similar to the last cycle and thus shift the last cycle by 11~years. } Additionally we consider as possible alternatives SSN patterns of half and twice its amplitude as shown in Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_f_plot.pdf}}
	\caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), that is, SIDC 13-month smoothed monthly actual SSN, SIDC prediction, SWPC prediction and by 11~years shifted SSN from previous cycle~24, together with two alternative trends of half and twice its amplitude.}
	\label{fig:SPP_orbit_predicted_SSN_overview_f_plot}
\end{figure}

@@

However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. \hl{The from direct measurements differing velocity and temperature extrapolations below about } \SI{20}{\Rs} \hl{indicate that PSP will indeed dive into the acceleration and heating regions of the solar wind to be expected at these distances } (see Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, \hl{because the region of the } Alfvénic \hl{critical surface, up to which the solar wind is believed to be accelerated, is predicted to lie in average around } \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000} \hl{and its position correlates with solar activity between } \SI{15}{\Rs} \hl{in cycle minimum and } \SI{30}{\Rs} in maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

>>>>

%\subsection{SSN fitting}
\textbf{Next we create solar activity dependent analytical representations of the solar wind frequency distributions. This is achieved by shifting the median positions of the lognormal distributions as a linear function of SSN. To enable these shifts }, we add a linear SSN dependency to the median
\begin{align}
	x_\text{med}(ssn) &= a_\text{med} \cdot ssn + b_\text{med}\,,	\label{eq:median_with_ssn}
\end{align}
using a factor to the SSN $a_\text{med}$ with a baseline $b_\text{med}$. We relate the mean with a scaling factor to the median to transfer its SSN dependency:
\begin{align}
	x_\text{avg}(ssn) &= (1 + a_\text{avg}) \cdot x_\text{med}(ssn)\,.	\label{eq:mean_with_ssn}
\end{align}
\textbf{These relations, substituted into the lognormal function (\ref{eq:single_lognormal_fit_function}), lead to a new SSN-dependent function $W'(x,ssn)$. This function} is then fitted to the yearly data\textbf{, using the yearly SSN as input parameter.} \textbf{The SSN is offset with the individual time lags for each parameter determined before, to profit from the higher correlation.} The three resulting fit coefficients ($a_\text{med}, b_\text{med}$ and $a_\text{avg}$) are presented in Table~\ref{tab:ssn_fit_parameters}.
\begin{table*}
	\caption{Resulting fit coefficients from the OMNI data fitting with \textbf{the linear SSN dependencies (\ref{eq:median_with_ssn}) and (\ref{eq:mean_with_ssn})}. For the velocity the fit parameters from the double lognormal fit \textbf{(\ref{eq:double_lognormal_fit_function})} and their balancing function \textbf{(\ref{eq:balance_with_ssn})} are given. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value. The listed SSN time lags are used for the fits.}}
	\label{tab:ssn_fit_parameters}
	\centering
	\begin{tabular}{l@{} c@{}
		S[table-format = 1.3(2)e+1]
		S[table-format = 1.4(2)]
		S[table-format = 1.3(2)e+1]
		S[table-format = +1.3(2)e+1]
		c c
		}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{2}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean}	&\multicolumn{2}{c}{Balance}	&\multicolumn{1}{c}{SSN lag}\\
		\cline{3-4}\cline{6-7}
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{SSN factor $a_\text{med}$}	&\multicolumn{1}{c}{Baseline $b_\text{med}$}	&\multicolumn{1}{c}{Scaling factor $a_\text{avg}$}	&\multicolumn{1}{c}{SSN factor $c_a$}	&\multicolumn{1}{c}{Baseline $c_b$}	&\multicolumn{1}{c}{[years]}\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&1.309(19)e-2	&4.285(17)	&8.786(78)e-2	&\multicolumn{1}{c}{--}	&--	&0\\
%		\multicolumn{2}{l}{Velocity}	&2.514(93)e-3	&3.8521(90)	&2.392(26)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Density}	&3.81(25)e-3	&4.495(26)	&3.050(27)e-1	&\multicolumn{1}{c}{--}	&--	&6\\
		\multicolumn{2}{l}{Temperature}	&1.974(26)e-2	&5.729(19)	&6.541(28)e-1	&\multicolumn{1}{c}{--}	&--	&3\\
		\hline
		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W'_1$}	&\multicolumn{1}{c}{--}	&3.633(12)	&1.008(37)e-2	&\multirow{2}{*}{\tablenum{-1.799(95)e-3}}	&\multirow{2}{*}{0.638(32)}	&\multirow{2}{*}{3}\\
			&\multicolumn{1}{c}{$W'_2$}	&\multicolumn{1}{c}{--}	&4.831(81)	&2.31(20)e-2	&	&	&\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{\hl{In units of } \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

@@

Next we retrieve the frequency distributions of the four \textbf{solar wind parameters in solar} distance bins of \SI{0.01}{\au}, choosing the same resolution as for the OMNI data analyzed in Sect.~\ref{sec:frequency_distribution}---the distributions \textbf{and their median values} are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}.
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Frequency distributions of the four solar wind parameters with respect to the solar distance. Plotted are the binned Helios data and the power law lognormal fit models with their median values (white lines). The double lognormal model is used for the velocity, its slow and fast parts are indicated by dotted lines.}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}

@@

For simplification, as mentioned before, we treat the exponents of the median and mean fit functions as being identical\textbf{, using one fit parameter for both.} Implementing the power law distance dependency~(\ref{eq:power_function}) into the lognormal function (\ref{eq:single_lognormal_fit_function}), we get the fit parameters ($d'_\text{med}$, $d'_\text{avg}$ and \textbf{the} common exponent $e'$). Again, we use the double lognormal function~(\ref{eq:double_lognormal_fit_function}) for the velocity distribution fit, resulting in $W''_\text{II}(x,r)$. The additional fit parameters are the balancing parameter $c'$ and for the second lognormal part $d'_\text{med,2}$ and $d'_\text{avg,2}$. The resulting fit coefficients for the four solar wind parameters are presented in Table~\ref{tab:extrapolation_model_fit_parameters}.

@@

\textbf{The power law lognormal models and the power law double lognormal model for the velocity, which result from the fitting, are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot} together with their median values.} The model’s magnetic field strength is broader around values of \SI{40}{nT} at the lower distance boundary than the data's frequency distribution implies. This behavior is expected because of the distance independent shape approximation applied. The velocity and temperature models’ upper values generally show a higher abundance than the actual data, see also zoom boxes in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}. \textbf{The increasing high velocity tail with distance} comes from using the same exponent for both slow and fast components. This effect is not seen in the data, more specifically, not only the slowest wind but also the fastest wind is expected to converge to more average speeds \citep{Sanchez-Diaz2016}.
%\citep{Sanchez-Diaz2016} p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction)

@@

%\subsection{Near-Sun extrapolation for PSP orbit}
%PSP orbit
Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching already 93~days after launch in November 2018 a first perihelion with a distance of \SI{0.16}{\au}. Seven additional Venus flybys allow \textbf{the perihelion distance to be reduced} to a minimum of \SI{9.86}{\Rs} \citep{Fox2015} as plotted in \textbf{the top panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_f_plot.pdf}}
	\caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), that is, SIDC 13-month smoothed monthly actual SSN, SIDC prediction, SWPC prediction and by 11~years shifted SSN from previous cycle~24, together with two alternative trends of half and twice its amplitude.}
	\label{fig:SPP_orbit_predicted_SSN_overview_f_plot}
\end{figure}

We extrapolate the derived empirical solar wind models (\ref{eq:general_sw_model}) to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}.

@@

	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. \textbf{The models obtained from Helios and OMNI measurements} are extrapolated to the PSP region---for the range between the extreme cases of solar minimum and maximum, i.e., $0 \le ssn \le 200$\textbf{. The lower edges of the shaded areas correspond to solar minimum, the upper edges to maximum.} Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \textbf{\citet{Sheeley1997} and \citet{Wang2000}}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}

@@

\textbf{The model and the extrapolation are visualized for the SSN range between solar minimum and maximum ($0 \le ssn \le 200$), indicated by the shaded regions in the figure.}

@@

The magnetic field strength is found to increase from median values of about \SI{43}{\nT} at \SI{0.25}{\au} to \SI{715}{\nT} at \SI{0.046}{\au} for a SSN of 0. Taking a SSN of 200 increases the values to \SI{69}{\nT} and \SI{1152}{\nT}. Our extrapolation results are slightly flatter than those derived from the analytical magnetic field model by \citet{Banaszkiewicz1998}, who constructed an analytic dipole plus quadrupole plus current sheet (DQCS) model for solar minimum.

@@

The average velocity is found to decrease from \SI{340}{\km\per\s} at \SI{0.25}{\au} to about \SI{290}{\km\per\s} at \SI{0.046}{\au} \textbf{3~years after a} SSN of 0 \textbf{occurred, whereas} using a SSN of 200 it decreases from \SI{390}{\km\per\s} to \SI{330}{\km\per\s}. Comparing the results with the measurements by \citet{Sheeley1997} and \citet{Wang2000} shows an overestimation in our extrapolated slow solar wind velocity values for distances below about \SI{20}{\Rs}. They used LASCO coronagraph observations to track moving coronal features (blobs) in the distance range \SIrange{2}{30}{\Rs} to determine speed profiles and sources of the slow solar wind and they derived temperature and sonic point values for slow solar wind with the isothermal expansion model from \citet{Parker1958}. Therefore, it generally can be expected that PSP will encounter a slower solar wind environment close to the Sun than our model estimates\textbf{. Thus} PSP will measure solar wind acceleration processes \citep{McComas2008}, maybe even still at \SI{30}{\Rs} as the study by \citet{Sheeley1997} suggests.

@@

The proton density increases from about \SI{84}{\per\cm\cubed} at \SI{0.25}{\au} to about \SI{3018}{\per\cm\cubed} at \SI{0.046}{\au} \textbf{6~years after a SSN of 0 occurred}. Being almost independent of the SSN the values for a SSN of 200 are only \SI{17}{\%} larger. The results are in good agreement with those of \citet{Leblanc1998}, who derived an electron density model from type~III radio burst observations. Their model shows that the density distance dependency scales with $r^{-2}$ and steepens just below \SI{10}{\Rs} with $r^{-6}$. \textbf{We assumed a solar wind helium abundance of \SI{5}{\%} to convert their electron densities to proton densities.}

@@

The extrapolated proton temperature increases from about \SI{260000}{\K} at \SI{0.25}{\au} to about \SI{1690000}{\K} at \SI{0.046}{\au} \textbf{3~years after a SSN of 0 occurred} and from \SI{440000}{\K} to \SI{2860000}{\K} for a SSN of 200. Knowing that near-Sun coronal temperatures are in the range of \SIrange{2}{3}{\mega\K} \citep{Billings1959,Liebenberg1975}, the model \textbf{overestimates} the extrapolated temperatures at the PSP perihelion distance.

%\subsection{SSN prediction for PSP mission time}
%SSN prediction
\hl{For SSN short-term predictions } several sources are available. The SIDC provides 12-month SSN forecasts\footnote{\url{http://sidc.be/silso/forecasts}} obtained from different methods (e.g., Kalman filter combined method). The SSN prediction of NOAA's Space Weather Prediction Center (SWPC) follows for the time period until end of 2019 a consensus of the Solar~Cycle~24 Prediction~Panel\footnote{\url{http://www.swpc.noaa.gov/products/solar-cycle-progression}}.
The SSN for PSP's first perihelion will be small---certainly below 20, whereas \textbf{the SSN during the closest perihelia, which will commence at the end of 2024 at the likely maximum phase of cycle~25, cannot be predicted at this time.}
\hl{For the prediction of the next solar cycle we simply assume a pattern similar to the last cycle and thus shift the last cycle by 11~years. } Additionally we consider as possible alternatives SSN patterns of half and twice its amplitude as shown in \textbf{the bottom panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.}

@@

However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. \textbf{The extrapolation results of the velocity and the temperature differ from the direct measurements of the other studies. This suggests that below about \SI{20}{\Rs} PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur} (see Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the \textbf{solar wind is believed to be accelerated up to the height of the Alfvénic critical surface, which is predicted to lie on average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000}, scaling with solar activity within a range of between \SI{15}{\Rs} at solar minimum and \SI{30}{\Rs} at solar} maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).
